Binning Strategies and Related Loss for Binned Scatterplots of Large Data

### 2016-12-03 versionForJCGS/PaperJCGS.Rnw
Fixed random binning net spatial loss problem from update to Figure 6.

# NEED TO SAVE TO BinningLossDemos.Rdata which is read in chunk 1
losses <- data.frame( bintype=NA, binsize=NA, type=NA, record=NA,
                      avg_loss=NA, avg_standardized_loss=NA,
                      avg_loss_x=NA,avg_loss_y=NA)
counter=0

for(type in "Uniform"){
  for(record in "Coarse"){
    for(binsize in seq(.5,8.5,by=.05)){
      counter = counter + 1
      alldatsub <- alldat[which(alldat$type==type & alldat$record==record),]
      reduced_data <- rect_bin_2d(data=alldatsub, xcol="x", ycol="y",originx=-1,originy=-1,
                                  widthx=binsize,widthy=binsize, output="reduced")
      losses[counter,] <- c("Standard",binsize,type,record, bin_loss(reduced_data))
      counter = counter + 1
      reduced_data <- rand_rect_bin_2d(data=alldatsub, xcol="x", ycol="y",originx=-1,originy=-1,
                                  widthx=binsize,widthy=binsize, output="reduced")
      losses[counter,] <- c("Random",binsize,type,record, bin_loss(reduced_data))
    }
  }
}

ggplot()+ 
  geom_line(aes(x=as.numeric(binsize), y=as.numeric(avg_loss), 
                color=bintype,group=bintype), data=losses,size=.5) +
  geom_smooth(aes(x=as.numeric(binsize), y=as.numeric(avg_loss), 
                  color=bintype,group=bintype),size=.25, alpha=.2,
              se=FALSE, data=losses,method="lm")+
  geom_point(aes(x=as.numeric(binsize), y=as.numeric(avg_loss), 
                 color=bintype,group=bintype),
             size=2,data=losses[losses$binsize %in% c(1,4,5), ])+
#   geom_text(aes(x=as.numeric(binsize), y=as.numeric(avg_loss), 
#                 label=paste(binsize,"X",binsize,sep=""),color=bintype,group=bintype),
#             size=10,data=losses[losses$binsize %in% c(1,4,5), ], 
#             nudge_x=rep())
  theme_bw() +
  scale_color_manual("",values=c("red","black"))+
  scale_x_continuous("Bin Sizes", breaks=1:8, labels=paste(1:8,"X",1:8,sep=""))+
  annotate("text", x=c(3, 6.5), y=c(2.5,1.75), label=c("Random","Standard"),
           color=c("red","black"),size=6)+
  theme(legend.position="none", 
         panel.grid.minor.x=element_blank()) + ylab("Average Spatial Loss")

@@

<<OriginOffsetSimulations, echo=F,include=T,eval=T,fig.width=8, fig.height=4, out.width='.9\\linewidth', fig.pos='h',fig.align='center',tidy=F, cache=TRUE, fig.show='hold',fig.cap="Net spatial loss 2X2 unit resolution data using various sized square bins over the range of possible origin offsets. The vertical gray lines indicate origin offset of (1,1).">>= 
##  Check out the spatial loss for the data that is recorded to resolution = alpha at various origin offsets
# standard and random binning with bin sizes that are scalar multiples of the data resolution
# save spatial loss information
# spatlossdatRound <- data.frame(NULL)
# index=0
# distindex = 0
# distrib <- c("Uniform", "Normal", "Exponential")
# widths <- seq(alpha,5*alpha,by=alpha)
# for (dat in 1:3){
#   distindex <- distindex + 1
#   for (width in widths){
#     for (origin in seq(-width,0, by=width/60)){
#       index <- index+1
#       binout <- RectBin2d(datlist[[dat]]$roundx,datlist[[dat]]$roundy,origin,origin,width,width,type="standard")
#       binout[[2]]$bintype <- "standard"
#       binout[[2]]$distrib <- distrib[distindex]
#       spatlossdatRound <- rbind(spatlossdatRound, binout[[2]])
#       index <- index+1
#       binout <- RectBin2d(datlist[[dat]]$roundx,datlist[[dat]]$roundy,origin,origin,width,width,type="random")
#       binout[[2]]$bintype <- "random"
#       binout[[2]]$distrib <- distrib[distindex]
#       spatlossdatRound <- rbind(spatlossdatRound, binout[[2]])
#       print(c(distrib[distindex],width,origin))
#     } 
#   }
# }
# write.csv(spatlossdatRound,"LossesByOriginBinwidth.csv", row.names=FALSE)
# lossOrigin <- read.csv("./data/LossesByOriginBinwidth.csv",header=TRUE)
# Save the loss binning data objects into central Rdata file for simplicity
# save(lossOrigin, spatlossdatSquareBins, allbindat, file="BinningLossDemos.Rdata")

>>>>

#   
# lossesUnif <- data.frame( bintype=NA, binsize=NA, type=NA, record=NA,
#                       avg_loss=NA, avg_standardized_loss=NA,
#                       avg_loss_x=NA,avg_loss_y=NA)
# counter=0
# 
# for(type in "Uniform"){
#   for(record in "Coarse"){
#     for(binsize in seq(.5,8.5,by=.05)){
#       counter = counter + 1 
#       alldatsub <- alldat[which(alldat$type==type & alldat$record==record),]
#       reduced_data <- rect_bin_2d(data=alldatsub, xcol="x", ycol="y",originx=-1,originy=-1,
#                                   widthx=binsize,widthy=binsize, output="reduced")
#       lossesUnif[counter,] <- c("Standard",binsize,type,record, bin_loss(reduced_data))
#       counter = counter + 1
#       reduced_data <- rand_rect_bin_2d(data=alldatsub, xcol="x", ycol="y",originx=-1,originy=-1,
#                                   widthx=binsize,widthy=binsize, output="reduced")
#       lossesUnif[counter,] <- c("Random",binsize,type,record, bin_loss(reduced_data))
#     }
#   }
# }
# # lossesUnif saved to BinningLossDemos.Rdata which is loaded in startup chunck

ggplot()+ 
  geom_line(aes(x=as.numeric(binsize), y=as.numeric(avg_loss), 
                color=bintype,group=bintype), data=lossesUnif,size=.5) +
  geom_smooth(aes(x=as.numeric(binsize), y=as.numeric(avg_loss), 
                  color=bintype,group=bintype),size=.25, alpha=.2,
              se=FALSE, data=lossesUnif,method="lm")+
  geom_point(aes(x=as.numeric(binsize), y=as.numeric(avg_loss), 
                 color=bintype,group=bintype),
             size=2,data=lossesUnif[lossesUnif$binsize %in% c(1,4,5), ])+
#   geom_text(aes(x=as.numeric(binsize), y=as.numeric(avg_loss), 
#                 label=paste(binsize,"X",binsize,sep=""),color=bintype,group=bintype),
#             size=10,data=lossesUnif[lossesUnif$binsize %in% c(1,4,5), ], 
#             nudge_x=rep())
  theme_bw() +
  scale_color_manual("",values=c("red","black"))+
  scale_x_continuous("Bin Sizes", breaks=1:8, labels=paste(1:8,"X",1:8,sep=""))+
  annotate("text", x=c(3, 6.5), y=c(2.5,1.75), label=c("Random","Standard"),
           color=c("red","black"),size=6)+
  theme(legend.position="none", 
         panel.grid.minor.x=element_blank()) + ylab("Average Spatial Loss")

@@

<<OriginOffsetSimulations, echo=F,include=T,eval=T,fig.width=8, fig.height=4, out.width='.9\\linewidth', fig.pos='h',fig.align='center',tidy=F, cache=TRUE, fig.show='hold',fig.cap="Net spatial loss 2X2 unit resolution data using various sized square bins over the range of possible origin offsets. The vertical gray lines indicate origin offset of (1,1).">>= 
##  Check out the spatial loss for the data that is recorded to resolution = alpha at various origin offsets
# standard and random binning with bin sizes that are scalar multiples of the data resolution
# save spatial loss information
# spatlossdatRound <- data.frame(NULL)
# index=0
# distindex = 0
# distrib <- c("Uniform", "Normal", "Exponential")
# widths <- seq(alpha,5*alpha,by=alpha)
# for (dat in 1:3){
#   distindex <- distindex + 1
#   for (width in widths){
#     for (origin in seq(-width,0, by=width/60)){
#       index <- index+1
#       binout <- RectBin2d(datlist[[dat]]$roundx,datlist[[dat]]$roundy,origin,origin,width,width,type="standard")
#       binout[[2]]$bintype <- "standard"
#       binout[[2]]$distrib <- distrib[distindex]
#       spatlossdatRound <- rbind(spatlossdatRound, binout[[2]])
#       index <- index+1
#       binout <- RectBin2d(datlist[[dat]]$roundx,datlist[[dat]]$roundy,origin,origin,width,width,type="random")
#       binout[[2]]$bintype <- "random"
#       binout[[2]]$distrib <- distrib[distindex]
#       spatlossdatRound <- rbind(spatlossdatRound, binout[[2]])
#       print(c(distrib[distindex],width,origin))
#     } 
#   }
# }
# write.csv(spatlossdatRound,"LossesByOriginBinwidth.csv", row.names=FALSE)
# lossOrigin <- read.csv("./data/LossesByOriginBinwidth.csv",header=TRUE)
# Save the loss binning data objects into central Rdata file for simplicity
# save(lossOrigin, spatlossdatSquareBins, allbindat, lossesUnif, file="./data/BinningLossDemos.Rdata")
